---
title: "HW2 Data Wrangling"
author: "Ronae McLin rkm2147"
date: "9/21/2020"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1

 Mr. Trash Wheel data

- We will be importing the data, and then follow up by cleaning it.

```{r}
trash_df = readxl::read_excel("./trash_wheel.xlsx",                    range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )

```

We are next going to read and clean precipitation data for 2017/2018. 

```{r}
rain_2018 = readxl::read_excel("./trash_wheel.xlsx",                    sheet = "2018 Precipitation",
        skip = 1) %>% 
        janitor::clean_names() %>% 
        drop_na(month) %>% 
        mutate(year = 2018) %>% 
        relocate(year)
  
rain_2017 = readxl::read_excel("./trash_wheel.xlsx",                    sheet = "2017 Precipitation",
        skip = 1) %>% 
        janitor::clean_names() %>% 
        drop_na(month) %>% 
        mutate(year = 2017) %>% 
        relocate(year)
```

Changing the variable of year into a complete month name and then combining 2017 and 2018 into one table. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

rain_df = bind_rows(rain_2017, rain_2018)

rain_df =
	left_join(rain_df, month_df, by = "month") %>% 
```

# Problem 2

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r}
transit_df = read_csv("./transit.csv") %>% 
   janitor::clean_names()
```


# Problem 3


first we will bring in pols-month.csv

 replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
```{r}
pols_df = read_csv("./538_data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year","month","day")) 



```
need to convert character into numeric value
unemployment.csv
```{r}

```

snp.csv
```{r}

```

