---
title: "HW2 Data Wrangling"
author: "Ronae McLin rkm2147"
date: "9/21/2020"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1

Read and clean the Mr. Trash Wheel sheet

to omit non-data entries (rows with notes / figures; columns containing notes) 

use reasonable variable names

omit rows that do not include dumpster-specific data

round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

```{r}
trash_df = readxl::read_excel("./trash_wheel.xlsx",                    range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    plastic_bottles:sports_balls,
    names_to = "items_collected",
    values_to = "number_items"
  )
```

okay, i'm thinking we should tidy up the month, year, date into one category

plastic bottles, polyst, trash bags, etc should go into one category called things collected


# Problem 2

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).
```{r}
transit_df = read_csv("./transit.csv") %>% 
   janitor::clean_names()
```


# Problem 3


first we will bring in pols-month.csv

 replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
```{r}
pols_df = read_csv("./538_data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year","month","day")) 



```
need to convert character into numeric value
unemployment.csv
```{r}

```

snp.csv
```{r}

```

